Behavior Driven Development in Haskell

### 2012-06-02 paper.tex
Related work: test-framework
We conclude

When wading through the numerous libraries written by
the Haskell community, one sometimes gets the impression that the type
system is being too much relied on, at the expense of testing.
While the type system cannot always guarantee semantic correctness,
test code has more applications than getting the code correct.
Test code assists developers in maintaining correctness when extending
software, and shows users how an API is supposed to be used.
To ease the production and enhance the value of test code, we advocate
behavior driven development with two tools:
Doctest integrated with QuickCheck for user documentation and
Hspec with new mnemonic combinators for code documentation.
Our method utilizes specification that can also be used for
documentation and automated testing. We hope that this three-in-one
quality motivates Haskellers to write more test code.

@@

%%%% Describe the problem.

@@

\noindent Our approach can be used for design, documentation and
automated testing.  We hope that this motivates Haskellers to write
more and better tests.

@@

\subsection{Applying BDD to Haskell: An Example}
This section shows how behavior driven development works in practice,
by means of the Fibonacci function.  Of course this is a very simple
example, but the method is also applicable to more sophisticated
situations where design decisions might not be as obvious as here.

@@

The {\tt test-framework} family is a TDD tool, which currently
integrates HUnit, QuickCheck, SmallCheck, and Doctest.
There is support for automatic test discovery that relies on naming
conventions for test functions\footnote{%
http://hackage.haskell.org/package/test-framework-th}.  It is
implemented with Template Haskell~\cite{template-haskell}, and it only
works within a single source file.  Automatic discovery across
multiple source files is not possible with a mechanism that is based
on Template Haskell, because Template Haskell can not generate import
statements\footnote{%
See http://hackage.haskell.org/trac/ghc/ticket/1475 }.

@@

\begin{quote}
TBD:
    \begin{itemize}
        \item Add "best current practice"
        \item We showed how to reuse pieces of specifications
    \end{itemize}
\end{quote}

This paper discussed how to apply behavior driven development to
Haskell.  We proposed to use Doctest for user documentation and Hspec
for code documentation. We added two markups for
properties and examples to Haddock and implemented
Doctest to verify examples and test properties from Haddock comments.
We also extended a set of combinators for expectations for Hspec.
With these two tools, specification can also be
used for documentation and automatic testing.  We also described a
mechanis that can automatically discover all specs of a project.  We
think that our approach enhances the value of test code, and hope
that three-in-one aspect will motivate Haskell programmers to write
more and better test code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{TBD}

>>>>

When wading through the numerous libraries written by
the Haskell community, one sometimes gets the impression that the type
system is being too much relied on, at the expense of testing.
While the type system cannot always guarantee semantic correctness,
test code has more applications than getting the code correct.
Test code assists developers in maintaining correctness when extending
software, and shows users how an API is supposed to be used.
To ease the production and enhance the value of test code, we advocate
behavior driven development with two tools:
Doctest integrated with QuickCheck for user documentation and
Hspec with new mnemonic combinators for code documentation.
Our approach can be used for design, documentation and automated
testing.  We hope that this motivates Haskell programmers to write
more and better tests.

@@

%%%% Describe the problem.
\begin{itemize}
    \item ref to appendix
\end{itemize}

@@

\noindent Our approach can be used for design, documentation and
automated testing.  We hope that this motivates Haskell programmers to write
more and better tests.

@@

\subsection{Applying BDD to Haskell: An Example}
\label{sec:bcp-example}
This section shows how behavior driven development works in practice,
by means of the Fibonacci function.  Of course this is a very simple
example, but the method is also applicable to more sophisticated
situations where design decisions might not be as obvious as here.

@@

{\tt test-framework} is a library that allows TDD in Haskell, and it
has support for HUnit, QuickCheck and SmallCheck.
There is support for automatic test discovery that relies on naming
conventions for test functions\footnote{%
http://hackage.haskell.org/package/test-framework-th}.  It is
implemented with Template Haskell~\cite{template-haskell}, and it only
works within a single source file.  Automatic discovery across
multiple source files is not possible with a mechanism that is based
on Template Haskell, because Template Haskell can not generate import
statements\footnote{%
See http://hackage.haskell.org/trac/ghc/ticket/1475 }.

@@

We showed how to apply behavior driven development to
Haskell.  We proposed to use Doctest for user documentation and Hspec
for code documentation. We extended Haddock's markup with support for
examples and properties.  And we implemented Doctest, which verifies
examples and tests properties in Haddock comments.
We added two missing features to Hspec: Combinators to express
expectations, and a mechanism for automatic discovery of specs.
And we showed how to reuse pieces of specification.
Our approach can be used for design, documentation and automated
testing.  We hope that this motivates Haskell programmers to write
more and better tests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{The final code from Section~\ref{sec:bcp-example}}
\label{sec:appendix-code}
